---
title: "Prep Validation Screen Data"
author: "Chinmaya"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(patchwork)
library(broom)
```

# Read in plate data

All the data from the plate reader is collected into a set of excel spreadsheets. The luminescence values show up at the bottom of the spreadsheet. I've manually matched up the names of the excel files to the appropriate cell line names and compounds in each plate. There is a text file with the same information in the same data folder I've read these files from.

The plate is organized with the concentrations decreasing over each column of the plate, following the concentration used in Klaeger. The last two columns in each plate are DMSO controls.

```{r}
read_csv_plate_data <- function(file_name,these_compounds,this_cell_line) {
	
	temp_plate = read_csv(file_name, col_names = FALSE) %>%
		select(-X1, -X12) %>%
		slice(2:7) %>%
		mutate(row = 1:6) %>% 
		mutate(compound = c(rep(these_compounds[1],3),rep(these_compounds[2],3))) %>%
		pivot_longer(-c(compound,row), names_to = "concentration_M",values_to = "lum") %>%
		mutate(cell_line = this_cell_line, plate_name = basename(file_name)) %>%
		mutate(lum = as.numeric(lum)) %>% 
		identity()
	
	DMSO_vals = temp_plate %>%
		filter(concentration_M == "X10" | concentration_M == "X11") %>%
		mutate(concentration_M = 0) %>%
		mutate(compound = "DMSO")
	
	temp_plate = temp_plate %>%
		filter(concentration_M != "X10", concentration_M != "X11") %>%
		mutate(concentration_M = case_when(
			concentration_M == "X2" ~ 3e-5,
			concentration_M == "X3" ~ 3e-6,
			concentration_M == "X4" ~ 1e-6,
			concentration_M == "X5" ~ 3e-7,
			concentration_M == "X6" ~ 1e-7,
			concentration_M == "X7" ~ 3e-8,
			concentration_M == "X8" ~ 1e-8,
			concentration_M == "X9" ~ 3e-9,
			T ~ NA_real_
		)) %>%
		bind_rows(DMSO_vals)
	
	return(temp_plate)
}
plate_vals = bind_rows(
	read_csv_plate_data(here('data/PRISM_validation_screen/BT_1.csv'),c("Sotrastaurin","ONO-4059 (analog)"),"BT-474"),
	read_csv_plate_data(here('data/PRISM_validation_screen/BT_2.csv'),c("MLN-2480","TAK-733"),"BT-474"),
	read_csv_plate_data(here('data/PRISM_validation_screen/BT_3.csv'),c("GDC-0994", "Ro-5126766"),"BT-474"),
	read_csv_plate_data(here('data/PRISM_validation_screen/BT_4.csv'),c("Lestaurtinib","UCN-01"),"BT-474"),
	read_csv_plate_data(here('data/PRISM_validation_screen/BT_5.csv'),c("PF-3758309","RGB-286638"),"BT-474"),
	# read_csv_plate_data(here('data/PRISM_validation_screen/BT_6.csv'),c("Pexmetinib", "K-252a"),"BT-474"),
	read_csv_plate_data(here('data/PRISM_validation_screen/BT_7.csv'),c("Pexmetinib", "K-252a"),"BT-474"),
	read_csv_plate_data(here('data/PRISM_validation_screen/BT_8.csv'),c("Sotrastaurin","TAK-733"),"BT-474"),
	
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_1.csv'),c("Sotrastaurin","ONO-4059 (analog)"),"HCC1806"),
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_2.csv'),c("MLN-2480","TAK-733"),"HCC1806"),
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_3.csv'),c("GDC-0994","Ro-5126766"),"HCC1806"),
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_4.csv'),c("Lestaurtinib","UCN-01"),"HCC1806"),
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_5.csv'),c("PF-3758309","RGB-286638"),"HCC1806"),
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_6.csv'),c("Pexmetinib", "K-252a"),"HCC1806"),
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_7.csv'),c("Pexmetinib", "K-252a"),"HCC1806"),
	read_csv_plate_data(here('data/PRISM_validation_screen/HCC_8.csv'),c("Sotrastaurin","TAK-733"),"HCC1806")
)

plate_order = c("Sotrastaurin","ONO-4059 (analog)","MLN-2480","TAK-733","GDC-0994",
								"Ro-5126766", "Lestaurtinib", "UCN-01","PF-3758309","RGB-286638","Pexmetinib",
								"K-252a","Pexmetinib", "K-252a", "Sotrastaurin", "TAK-733")

```

# Per Plate DMSO Boxplots

```{r}
all_DMSO_vals = plate_vals %>%
	filter(compound == "DMSO") %>%
	mutate(plate_name = paste0(plate_name))

ggplot(all_DMSO_vals, aes(x=plate_name,y=lum)) + 
	geom_boxplot() +
	# ylim(c(min(plate_vals$lum), NA)) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave(here('figures/validation_screen/DMSO_boxplots.png'))
```

# Per Plate Diagnostic Plotting/Analysis

```{r}
# DMSO_checks = plate_vals %>% 
# 	filter(compound == "DMSO") %>%
# 	group_by(plate_name) %>%
# 	nest() %>%
# 	mutate(DMSO_1 = map_dbl(data, ~ tidy(t.test(lum ~ target_row, 
# 																							data = .x %>% 
# 																								mutate(target_row = ifelse(row == 1, 1, 0))))$p.value)) %>%
# 	mutate(DMSO_2 = map_dbl(data, ~ tidy(t.test(lum ~ target_row, 
# 																							data = .x %>% 
# 																								mutate(target_row = ifelse(row == 2, 1, 0))))$p.value)) %>%
# 	mutate(DMSO_3 = map_dbl(data, ~ tidy(t.test(lum ~ target_row, 
# 																							data = .x %>% 
# 																								mutate(target_row = ifelse(row == 3, 1, 0))))$p.value)) %>%
# 	mutate(DMSO_4 = map_dbl(data, ~ tidy(t.test(lum ~ target_row, 
# 																							data = .x %>% 
# 																								mutate(target_row = ifelse(row == 4, 1, 0))))$p.value)) %>%
# 	mutate(DMSO_5 = map_dbl(data, ~ tidy(t.test(lum ~ target_row, 
# 																							data = .x %>% 
# 																								mutate(target_row = ifelse(row == 5, 1, 0))))$p.value)) %>%
# 	mutate(DMSO_6 = map_dbl(data, ~ tidy(t.test(lum ~ target_row, 
# 																							data = .x %>% 
# 																								mutate(target_row = ifelse(row == 6, 1, 0))))$p.value))
# 	select(-data) %>%
# 	pivot_longer(-plate_name, values_to = "p.value", names_to = "DMSO_row") %>%
# 	mutate(p.value.adj = p.adjust(p.value, method="fdr"))
```


```{r}
treatment_lum_summaries = plate_vals %>%
	filter(compound != "DMSO") %>% 
	group_by(compound, concentration_M, cell_line) %>%
	summarise(sd = sd(lum),
						num_samples = n(),
						mean_lum = mean(lum),
						CV = sd/mean_lum)

ggplot(treatment_lum_summaries, aes(x=CV)) + geom_histogram()
ggsave(here('figures/validation_screen/lum_distribution.png'))

```

# Per row Normalization

```{r}
plate_DMSO_means = plate_vals %>%
	filter(compound == "DMSO") %>%
	group_by(row, plate_name, cell_line) %>%
	summarise(DMSO_mean = mean(lum),
						num_DMSO = n())

if (! all(plate_DMSO_means$num_DMSO == max(plate_DMSO_means$num_DMSO))) {
	stop("Number of DMSO values in a plate isn't correct.")
} else {
	plate_DMSO_means = plate_DMSO_means %>% select(-num_DMSO)
}

```

```{r}
qc_filter <- function(data_set, viability_threshold = 125) {
	qc_fails = data_set %>% 
		filter(viability > viability_threshold) %>% 
		select(row,compound,cell_line,plate_name) %>% 
		unique() %>%
		mutate(qc_fail = T)
	
	data_set = data_set %>%
		left_join(qc_fails) %>%
		mutate(qc_fail = ifelse(is.na(qc_fail), F, T))
	
	return(data_set)
}

dir.create(here('results/validation_results/'), showWarnings = F)
plate_vals_row_norm = plate_vals %>%
	left_join(plate_DMSO_means) %>% 
	mutate(viability = lum/DMSO_mean*100) %>%
	qc_filter() %>%
	write_rds(here('results/validation_results/validation_screen.rds'))

treatment_via_summaries = plate_vals_row_norm %>% 
	group_by(compound, concentration_M, cell_line) %>%
	summarise(sd = sd(viability),
						num_samples = n(),
						mean_via = mean(viability),
						CV = sd/mean_via)

```

```{r}
ggplot(plate_vals_row_norm %>% filter(compound != "DMSO"), aes(x=concentration_M, y = viability, color=qc_fail)) +
	geom_hline(aes(yintercept = 90),alpha=0.5) +
	geom_hline(aes(yintercept = 100),alpha=0.5, linetype = 2) +
	geom_hline(aes(yintercept = 125),alpha=0.5, linetype = 2) +
	scale_x_log10() +
	geom_jitter() +
	geom_smooth() +
	BerginskiRMisc::theme_berginski() +
	facet_grid(rows = vars(cell_line), cols = vars(compound)) +
	ggtitle("")

ggsave(here('figures/validation_screen/validation_viabilities_row_norm.png'),width=20*0.75,height=15*0.75)
BerginskiRMisc::trimImage(here('figures/validation_screen/validation_viabilities_row_norm.png'))
```